Cluster Difference Imaging Photometric Survey. III. An Open Cluster Spread Across 500\,pc

### 2021-03-12 cluster_paper/ms.tex
update ms

\title{
  Cluster Difference Imaging Photometric Survey. III.
  %The Coeval Core \& Halo of NGC\,2516
  %The Halo of NGC\,2516 Extends To At Least 300\,pc
  %An Open Cluster With a 300\,pc Halo
  An Open Cluster Spread Across 400\,pc
  %NGC\,2516  300\,pc
}

@@

% NB. the "classical tidal radius" is from e.g., Piskunov+08 or
% Kharchenko+2013. The latter givs a radius r2 of 0.71 degrees. At 380
% pc, corresponds to 4.7pc. The King tidal radius given by K13 is
% 7.70pc.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these diffuse halos could be tidal escapees from the
  cluster cores, or they could be remnants of an initially dispersed
  star formation complex.
  They could also be field stars---the false positives of overly
  sensitive clustering algorithms.
  In this study, we focus on a halo with a tip-to-tip length of
  $\approx$400\,pc reported around the $\approx$150\,Myr open cluster
  NGC\,2516.
  The classical tidal radius of the cluster is $\lesssim$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two in three kinematically selected halo members 25-250\,pc from the
  cluster center show rotation periods consistent with a
  gyrochronological age of 150\,Myr.
  A comparison sample of field stars shows no such trend.
  The rotation period and lithium abundances of the cluster stars are
  also anti-correlated at fixed stellar mass, as has been observed in
  similar stellar populations.
  This work quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants, and
  expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

A separate project that benefits from the new discoveries of dispersed
stellar populations is that of finding young transiting exoplanets.
Young transiting planets are hard to find because young stars are
rare, and mostly reside in crowded galactic plane \citep[see {\it
e.g.},][]{Kharchenko_et_al_2013,piskunov_global_2018}.  If the
dispersed halos of nearby star clusters could be reliably identified,
this could expand the census of nearby young stars by up to a factor
of $10$, based on the expected fraction of stars thought to be lost
during gas dispersal.

@@

One point of difficulty however is that different clustering methods
yield differing levels of both sensitivity and precision
\citep{hunt_clustering_2020}.  Using say a Gaussian Mixture Model
tautologically yields open clusters that are elliptical \citep[{\it
e.g.},][]{wallace_m4_2018}.  Some unsupervised methods yield dispersed
and asymmetric structures with number densities down to 100 times less
than the field around the same regions (see {\it e.g.}
\citealt{kounkel_untangling_2019} and \citealt{meingast_2021}).
Differences between different clustering methods raise the issue of
how contaminated the resulting samples are, and in some cases whether
the reported structures are truly coeval stellar populations at all.

@@

%TODO: double check this owrding.
A brief note on terminology.  Low-density stellar associations
connected to a dense population (the ``core'') have been described as
being in ``halos'', ``strings'', ``coronae'', ``snakes'',
``outskirts'', and ``tidal tails'' \citep[{\it
e.g.},][]{davenport_death_2010,kounkel_untangling_2019,roser_hyades_2019,tian_discovery_2020,meingast_2021}.
The latter term implies a particular model for the formation of the
association.  We opt for ``halo'' because it is concise and model
agnostic.  It suffers however from the suggestion of spherical
symmetry.  The halo of NGC\,2516 might be better described as ``tails'',
but the primary goal of this study is not to verify whether these
tails are tidal.

@@

We selected candidate \cn\ members based on those reported in the
literature.  While a number of pre-Gaia studies have been performed,
the purity and accuracy of the Gaia-derived results are the current
state of the art \citep{jeffries_ngc2516_2001,Kharchenko_et_al_2013}.
We therefore adopted what we viewed as the most interesting clustering
samples to compare: those of \citet{cantatgaudin_gaia_2018},
\citet{kounkel_untangling_2019} and \citet{meingast_2021}.  We refer
to these studies in the following as
\citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}
respectively.  A useful visualization of the different samples is
available online\footnote{
  \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
  made by \citet{meingast_2021}.}.  While we considered performing our
own clustering analysis using the Gaia data, such an effort would in
effect only replicate the work of these investigators.  We opt to
instead use their studies as starting points.

The Gaia clustering studies each used different selection functions,
which yielded different results.  \citetalias{cantatgaudin_gaia_2018}
considered stars brighter than $G=18$\,mag within a few degrees of the
cluster center, and reported 1106 candidate members of \cn.
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
considered stars up to $\approx$1\,mag fainter, and reported 3003 and
1860 members respectively.  The unsupervised clustering techniques
that each of these studies applied to the second Gaia data release are
discussed Appendix~\ref{app:clustering}, as is the overlap between
their resulting membership samples.

@@

To quantify this, we compute the separation of the positions and 2D
velocities from the median \citetalias{cantatgaudin_gaia_2018} core
member.  The result is shown on the right side of
Figure~\ref{fig:physical_x_rotn}.  The halo extends to separations of
about 200\,pc in physical space from the cluster core.  This
corresponds to total a length of $\approx$400\,pc, depending on which
members of the halo are chosen as the ``tips'' on either end.  This
agrees broadly speaking with the structure of the halo reported by
\citetalias{kounkel_untangling_2019}.

>>>>

\title{
  Cluster Difference Imaging Photometric Survey. III.
  %The Coeval Core \& Halo of NGC\,2516
  %The Halo of NGC\,2516 Extends To At Least 300\,pc
  %An Open Cluster With a 300\,pc Halo
  An Open Cluster Spread Across 500\,pc
  %NGC\,2516  300\,pc
}

@@

% NB. the "classical tidal radius" is from e.g., Piskunov+08 or
% Kharchenko+2013. The latter givs a radius r2 of 0.71 degrees. At 380
% pc, corresponds to 4.7pc. The King tidal radius given by K13 is
% 7.70pc.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these diffuse halos could be tidal escapees from the
  cluster cores, or they could be remnants of an initially dispersed
  star formation complex.
  They could also be field stars---the false positives of overly
  sensitive clustering algorithms.
  In this study, we focus on a halo with a tip-to-tip length of
  $\approx$500\,pc reported around the $\approx$150\,Myr open cluster
  NGC\,2516.
  The classical tidal radius of the cluster is $\lesssim$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two in three kinematically selected halo members out to 250\,pc from the
  cluster center show rotation periods consistent with a
  gyrochronological age of 150\,Myr.
  A comparison sample of field stars shows no such trend.
  The rotation period and lithium abundances of the cluster stars are
  also anti-correlated at fixed stellar mass, as has been observed in
  similar stellar populations.
  This work quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants, and
  expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

A separate project that benefits from the new discoveries of dispersed
stellar populations is that of finding young transiting exoplanets.
Young transiting planets are hard to find because young stars are
rare, and mostly reside in the crowded galactic plane \citep[see {\it
e.g.},][]{Kharchenko_et_al_2013,piskunov_global_2018}.  If the
dispersed halos of nearby star clusters could be reliably identified,
this could expand the census of nearby young stars by up to a factor
of $10$, based on the expected fraction of stars thought to be lost
during gas dispersal.

@@

One point of difficulty however is that different clustering methods
yield differing levels of both sensitivity and precision
\citep{hunt_clustering_2020}.  Using say a Gaussian Mixture Model
tautologically yields open clusters that are elliptical \citep[{\it
e.g.},][]{wallace_m4_2018}.  Some unsupervised methods yield dispersed
and asymmetric structures with number densities down to 100 times less
than the field around the same regions (see {\it e.g.},
\citealt{kounkel_untangling_2019} and \citealt{meingast_2021}).
Differences between different clustering methods raise the issue of
how contaminated the resulting samples are, and in some cases whether
the reported structures are truly coeval stellar populations at all.

@@

%TODO: double check this owrding.
A brief note on terminology.  Low-density stellar associations
connected to a dense population (the ``core'') have been described as
being in ``halos'', ``strings'', ``coronae'', ``snakes'',
``outskirts'', and ``tidal tails'' \citep[{\it
e.g.},][]{davenport_death_2010,kounkel_untangling_2019,roser_hyades_2019,tian_discovery_2020,meingast_2021}.
The latter term implies a particular model for the formation of the
association.  We opt for ``halo'' because it is concise and model
agnostic.  It suffers however from the suggestion of spherical
symmetry.  The halo of NGC\,2516 might be best described as ``tails'',
but the primary goal of this study is not to verify whether these
tails are tidal.

@@

We selected candidate \cn\ members based on those reported in the
literature.  While some pre-Gaia analyses were available,
the purity and accuracy of the Gaia-derived results are the current
state of the art \citep{jeffries_ngc2516_2001,Kharchenko_et_al_2013}.
We therefore adopted what we viewed as the most interesting clustering
samples to compare: those of \citet{cantatgaudin_gaia_2018},
\citet{kounkel_untangling_2019} and \citet{meingast_2021}.  We refer
to these studies in the following as
\citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}
respectively.  A useful visualization of the different samples is
available online\footnote{
  \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
  made by \citet{meingast_2021}.}.  While we considered performing our
own clustering analysis using the Gaia data, such an effort would in
effect only replicate the work of these investigators.  We opt to
instead use their studies as starting points.

The Gaia clustering studies each used different selection functions,
which yielded different results.  \citetalias{cantatgaudin_gaia_2018}
considered stars brighter than $G=18$\,mag within a few degrees of the
cluster center, and reported 1106 candidate \cn\ members.
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
considered stars up to $\approx$1\,mag fainter, and reported 3003 and
1860 members respectively.  The unsupervised clustering techniques
that each of these studies applied to the second Gaia data release are
discussed Appendix~\ref{app:clustering}, as is the overlap between
their resulting membership samples.

@@

To quantify this, we compute the separation of the positions and 2D
velocities from the median \citetalias{cantatgaudin_gaia_2018} core
member.  The result is shown on the right side of
Figure~\ref{fig:physical_x_rotn}.  The halo extends to separations of
about 200\,pc in physical space from the cluster core.  This
corresponds to total a length of $\approx$500\,pc, depending on which
members of the halo are chosen as the ``tips'' on either end.  This
agrees broadly speaking with the structure of the halo reported by
\citetalias{kounkel_untangling_2019}.
